\textbf{{\Large {Recursive algorithms, branching coefficients affine algebras and applications}}}

### 2009-12-25 AffineLieAlgebras/InformalNotes/article.tex
Merge of physical example and notes of Vladimir Dmitrievich

\title{\textbf{{\Large {Recursive algorithms, branching coefficients affine algebras and applications}}}}
\author{Vladimir Lyakhovsky \thanks{ Supported by 
 RFFI grant N 09-01-00504 and the National Project RNP.2.1.1./1575 }\\
Theoretical Department, SPb State University,\\
198904, Sankt-Petersburg, Russia \\
e-mail:lyakh1507@nm.ru \\
[5mm] Anton Nazarov \thanks{ Supported by
the National Project RNP.2.1.1./1575 }\\
Theoretical Department, SPb State University,\\
198904, Sankt-Petersburg, Russia \\
e-mail:antonnaz@gmail.com
}
\maketitle

@@

\newpage
\section{Introduction}
\label{sec:introduction}

@@

The extended Dynkin diagram of $B_2$ is presented in the Figure 1.
\begin{figure}[ph]
  \centering
  \includegraphics[width=50mm]{B2_A1_diagram.pdf}
  \caption{Extended Dynkin diagram of $B_2$ and embedding of $A_1$}
  \label{fig:B2Dynkin}
\end{figure}
When we drop the central node and get the embedding $A_1\oplus A_1\subset B_2$. Then $\mathfrak{a}=A_1$ and $\mathfrak{a}_{\bot}=A_1$.

@@

Anomalous branching coefficient $k^{(1,0)}_{1}=2$, then for anomalous branching coefficient $k^{(1,0)}_{0}$ the formula (\ref{recurrent relation}) gives us
\begin{equation}
  \label{eq:23}
  k^{(1,0)}_{0}=-1\cdot k^{(1,0)}_2 +2\cdot k^{(1,0)}_1 - 3\cdot \delta_{0,0} = 1
\end{equation}
So we have computed the branching coefficients.
\newpage
\subsubsection{Embedding of $B_2$ into $B_4$}
\label{sec:someth-high-dimens}
Consider the the regular embedding of the subalgebra $B_2$ into the algebra $B_4$.
We calculate the branching coefficients for the fundamental representation of $B_4$.
The corresponding Dynkin diagrams are in the Figure \ref{fig:dynkin}.
\begin{figure}[ph]
  \centering
  \includegraphics[width=60mm]{B4_B2_2A1.pdf}
  \caption{Dynkin diagrams}
  \label{fig:dynkin}
\end{figure}

@@

We show the result of this procedure on the Figure \ref{fig:B4B2anom}.
\begin{figure}[h]
    \includegraphics[width=170mm]{B4_B2_anom_points.pdf}
  \caption{Anomalous points with the dimensions of corresponding $\mathfrak{a}_{\bot}$-modules.}
  \label{fig:B4B2anom}
\end{figure}

@@

Using the definition (\ref{fan-defined}) we get the following set of
the points $\Gamma$ with the corresponding values $s(\gamma+\gamma_0)$, depicted at the Figure \ref{fig:B4B2Fan}.
\begin{figure}[h]
  \includegraphics[width=150mm]{B4_B2_fan.pdf}
  \caption{Fan for $B_2\subset B_4$}
  \label{fig:B4B2Fan}
\end{figure}
We use the recurrent relation (\ref{recurrent relation}) and get
following branching coefficients:
\begin{equation}
  \label{eq:24}
  \pi_{\mathfrak{a}} \left(ch L^{(0,1,0,2)}_{B_4}\right) = 6 \; ch L^{(0,0)}_{B_2}+ 60
  \; ch L_{B_2}^{(0,2)}+ 30 \; ch L_{B_2}^{(1,0)}+ 19 \; ch L_{B_2}^{(2,0)}+
  40 \; ch L_{B_2}^{(1,2)}+ 10 \; ch L_{B_2}^{(2,2)}
\end{equation}
The dimension of the highest-weight $B_4$-module $L^{(0,1,0,2)}_{B_4}$
is equal to 2772. It is easy to see, that right-hand side of the
equation (\ref{eq:24}) gives the same result.

\newpage
\section{Physical applications}
\label{sec:phys-appl}
Here we want to discuss possible applications of the described techniques in physical models.

@@

\begin{enumerate}
\item Construct the set $\Delta^{+}$ of the positive roots of the algebra $\mathfrak{g}$.
\item Select the positive roots $\alpha\in \Delta^{+}$ which are orthogonal to the root subspace of the subalgebra $\mathfrak{a}$ and form the set $\Delta^{+}_{\bot}$.
\item Construct the set $\widehat{\Psi^{(\mu)}}=\left\{\omega(\mu+\rho)-\rho;\; \omega\in W\right\}$ of the anomalous points of the $\mathfrak{g}$-module $L^{(\mu)}$.
\item Select those weights $\lambda=\omega(\mu+\rho)$ which lies in the closure of the main Weyl chamber of the algebra $\mathfrak{a}_{\bot}$. Since we have constructed the set $\Delta^{+}_{\bot}$ we can easily check if the weight $\lambda$ lies in the main Weyl chamber of $\mathfrak{a}_{\bot}$ checking that the scalar product of $\lambda$ with the roots of $\Delta^{+}_{\bot}$ is non-negative.
\item For $\lambda=\omega(\mu+\rho),\; \lambda\in \bar{C}_{\mathfrak{a}_{\bot}}$ calculate the dimensions of the corresponding modules $\mathrm{dim}\left(L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}\right)$ using the Weyl formula with the set $\Delta^{+}_{\bot}$.
\item Construct the set $\Gamma$ \eqref{fan-defined}.
\item Calculate anomalous branching coefficients in the main Weyl
  chamber of the subalgebra $\mathfrak{a}$ using recurrent relation (\ref{recurrent relation}).
\end{enumerate}

There exist three level one dominant weights $[1,0,0],\;[0,1,0],\;[0,0,1]$ in the weight space of $\hat{A}_2$.
It is easy to see that the set $\Delta_{\bot}$ is trivial. 
\subsubsection{Special embedding $\hat{su}(2)\subset \hat{sp}(4)$.}
\label{sec:spec-embedd-wideh}

Consider the level ten embedding of the affine $\hat A_1$ into the level one affine $\hat C_2$. 
The relation (\ref{eq:33}) and the finite reducibility theorem greatly simplify the calculation with the algorithm \ref{sec:algorithm}, since we know that there exists only finite number of the non-zero branching coefficients and the maximal grade of this coefficients is restricted.

This affine embedding is the affine extension of the special embedding $A_1\subset C_2$ with the embedding index equal to 10. The projection of the weights of algebra $C_2$ is calculated using the projection matrix $\mathcal{P}=(3,4)$. The embedding index
\begin{equation}
  \label{eq:37}
  x_e=\frac{\left|\mathcal{P}\Theta_{\mathfrak{g}}\right|^2}{\left|\Theta_{\mathfrak{a}}\right|^2}=\frac{k_{\mathfrak{a}}}{k_{\mathfrak{g}}}=10
\end{equation}
If we consider this embedding in the orthogonal basis where the simple roots of $C_2$ are given by
\begin{equation}
  \label{eq:38}
  \alpha_1=(1,-1),\quad \alpha_2=(0,2),
\end{equation}
then the simple root of the embedded $A_1$ is equal to
\begin{equation}
  \label{eq:39}
  \beta=\sqrt{\frac{2}{5}}(3,1)
\end{equation}
It is easy to see that the set $\Delta^{+}_{\bot}$ of the $C_2$ roots orthogonal to $\beta$ is empty.

Now we proceed the computation of the branching coefficients using the algorithm \ref{sec:algorithm}. 
On the level 1 there exist three dominant weights with Dynkin labels $[1,0,0], \; [0,1,0],\; [0,0,1]$.

Do the calculation, show the restriction on the grade and the final result
  \begin{equation}
    \label{eq:35}
    Z=\left|\chi_{[10,0]}+\chi_{[4,6]}\right|^2+\left|\chi_{[7,3]}+\chi_{[3,7]}\right|^2+\left|\chi_{[6,4]}+\chi_{[0,10]}\right|^2
  \end{equation}

>>>>

\pagenumbering{arabic}
\title{\textbf{{\Large {Recursive algorithms, branching coefficients affine algebras and applications}}}}
\author{Vladimir Lyakhovsky \thanks{ Supported by 
 RFFI grant N 09-01-00504 and the National Project RNP.2.1.1./1575 }\\
Theoretical Department, SPb State University,\\
198904, Sankt-Petersburg, Russia \\
e-mail:lyakh1507@nm.ru \\
[5mm] Anton Nazarov \thanks{ Supported by
the National Project RNP.2.1.1./1575 }\\
Theoretical Department, SPb State University,\\
198904, Sankt-Petersburg, Russia \\
e-mail:antonnaz@gmail.com
}
\maketitle

@@

\section{Introduction}
\label{sec:introduction}

@@

The extended Dynkin diagram of $B_2$ is presented in the Figure \ref{fig:B2Dynkin}.
\begin{figure}[ph]
  \centering
  \includegraphics[width=50mm]{B2_A1_diagram.pdf}
  \caption{Extended Dynkin diagram of $B_2$ and embedding of $A_1$}
  \label{fig:B2Dynkin}
\end{figure}
When we drop the central node and get the embedding $A_1\oplus A_1\subset B_2$. Then $\mathfrak{a}=A_1$ and $\mathfrak{a}_{\bot}=A_1$.

@@

Anomalous branching coefficient $k^{(1,0)}_{1}=2$, then for anomalous branching coefficient $k^{(1,0)}_{0}$ the formula (\ref{recurrent relation}) gives us
\begin{equation}
  \label{eq:23}
  k^{(1,0)}_{0}=-1\cdot k^{(1,0)}_2 +2\cdot k^{(1,0)}_1 - 3\cdot \delta_{0,0} = 1
\end{equation}
So we have computed the branching coefficients.

\subsubsection{Embedding of $B_2$ into $B_4$}
\label{sec:someth-high-dimens}
Consider the the regular embedding of the subalgebra $B_2$ into the algebra $B_4$.
We calculate the branching coefficients for the fundamental representation of $B_4$.
The corresponding Dynkin diagrams are in the Figure \ref{fig:dynkin}.
\begin{figure}[ph]
  \centering
  \includegraphics[width=60mm]{B4_B2_2A1.pdf}
  \caption{Dynkin diagrams}
  \label{fig:dynkin}
\end{figure}

@@

We show the result of this procedure on the Figure \ref{fig:B4B2anom}.
\begin{figure}[h!tb]
    \includegraphics[width=130mm]{B4_B2_anom_points.pdf}
  \caption{Anomalous points with the dimensions of corresponding $\mathfrak{a}_{\bot}$-modules.}
  \label{fig:B4B2anom}
\end{figure}

@@

Using the definition (\ref{fan-defined}) we get the following set of
the points $\Gamma$ with the corresponding values $s(\gamma+\gamma_0)$, depicted at the Figure \ref{fig:B4B2Fan}.
\begin{figure}[h!tb]
  \includegraphics[width=130mm]{B4_B2_fan.pdf}
  \caption{Fan for $B_2\subset B_4$}
  \label{fig:B4B2Fan}
\end{figure}
We use the recurrent relation (\ref{recurrent relation}) and get
following branching coefficients:
\begin{equation}
  \label{eq:24}
  \pi_{\mathfrak{a}} \left(ch L^{(0,1,0,2)}_{B_4}\right) = 6 \; ch L^{(0,0)}_{B_2}+ 60
  \; ch L_{B_2}^{(0,2)}+ 30 \; ch L_{B_2}^{(1,0)}+ 19 \; ch L_{B_2}^{(2,0)}+
  40 \; ch L_{B_2}^{(1,2)}+ 10 \; ch L_{B_2}^{(2,2)}
\end{equation}
The dimension of the highest-weight $B_4$-module $L^{(0,1,0,2)}_{B_4}$
is equal to 2772. It is easy to see, that right-hand side of the
equation (\ref{eq:24}) gives the same result.

\subsection{Affine Lie algebras}
\label{sec:affine-lie-algebras}
\subsubsection{Embedding of the affine algebra into affine algebra}
\label{sec:embedd-affine-algebr}

Consider the affine extension of the example \ref{sec:regul-embedd-a_1}. 
Since this embedding is regular, the level of the representations of the subalgebra is equal to the level of the representation of the algebra.

The set $\Delta^{+}_{\bot}$ of the orthogonal positive roots with the zero projection on the root space of the subalgebra $\hat{A_1}$ is the same as in the finite-dimensional case.

Consider the level one representation of the algebra $\mathfrak{g}=\hat{B_2}$ with the highest weight $w_1=(1,0,1,0)$, where the first two components are the coordinates of the classical part in the orthogonal basis $e_1,e_2$, the third is is the level of the weight and the fourth the grade.

The set of the anomalous points of this representation up to sixth grade is depicted in the Figure \ref{fig:affine_B2_anom_point} and in each grade it looks like in the Figure \ref{fig:B2_A1}.

\begin{figure}[h!tb]
  \includegraphics[width=160mm]{AffineB2_A1_Anom.pdf}
  \caption{The anomalous points of the $(1,0,1,0)$ representation of the algebra $\hat{B_2}$}
  \label{fig:affine_B2_anom_point}
\end{figure}

As the next step of our algorithm \ref{sec:algorithm} we project the anomalous points to the weight space of the subalgebra $\hat{A_1}$ and calculate the dimensions of the corresponding $\mathfrak{a}_{\bot}$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$.
The result of this computation up to the twelfth grade is presented at the Figure
\begin{figure}[h!tb]
  \centering
  \includegraphics[width=150mm]{AffineB2_A1_proj_anom.pdf}
  \caption{Projected anomalous points and the dimensions of $\mathfrak{a}_{\bot}$-modules.}
  \label{fig:AffineB2_A1_anom_proj}
\end{figure}

Using the definition (\ref{fan-defined}) we get the following set of
the points $\Gamma$ with the corresponding values $s(\gamma+\gamma_0)$: \ref{fig:AffineB2A1Fan}.
Here we restricted the computation to the twelfth grade.
\begin{figure}[ph]
  \centering
  \includegraphics[width=130mm]{AffineB2_A1_fan.pdf}
  \caption{Fan for $\hat{A_1}\subset \hat{B_2}$}
  \label{fig:AffineB2A1Fan}
\end{figure}

Also we should mention that the lowest vector of the fan $\gamma_0$ is equal to zero, since we have excluded all the roots of $\Delta^{+}_{\bot}$ from the defining relation (\ref{fan-defined}).

Using the recurrent relation for the anomalous branching coefficients we get the following result
\begin{figure}[ph]
  \centering
  \includegraphics[width=130mm]{AffineB2_A1_branching.pdf}
  \caption{Anomalous branching coefficients for $\hat{A_1}\subset \hat{B_2}$}
  \label{fig:AffineB2_A1_branching}
\end{figure}

Selecting the points inside the main Weyl chamber of the subalgebra $\hat{A_1}$ we get the following results for the branching coefficients up to twelfth grade
\begin{multline}
  \label{eq:28}
  L^{w_1}_{\hat{B_2}\downarrow \hat{A_1}}=2 L_{\hat{A_1}}^{w_1}(0)\oplus 1 L_{\hat{A_1}}^{w_0}(0)\oplus 4 L_{\hat{A_1}}^{w_0}(-1)\oplus\\
    2 L_{\hat{A_1}}^{w_1}(-1)\oplus 8 L_{\hat{A_1}}^{w_0}(-2)\oplus
    8 L_{\hat{A_1}}^{w_1}(-2)\oplus 15 L_{\hat{A_1}}^{w_0}(-3)\oplus\\
    12 L_{\hat{A_1}}^{w_1}(-3)\oplus 26 L_{\hat{A_1}}^{w_1}(-4)\oplus
    29 L_{\hat{A_1}}^{w_0}(-4)\oplus 51 L_{\hat{A_1}}^{w_0}(-5)\oplus\\
    42 L_{\hat{A_1}}^{w_1}(-5)\oplus 78 L_{\hat{A_1}}^{w_1}(-6)\oplus
    85 L_{\hat{A_1}}^{w_0}(-6)\oplus 120 L_{\hat{A_1}}^{w_1}(-7)\oplus\\
    139 L_{\hat{A_1}}^{w_0}(-7)\oplus 202 L_{\hat{A_1}}^{w_1}(-8)\oplus
    222 L_{\hat{A_1}}^{w_0}(-8)\oplus 306 L_{\hat{A_1}}^{w_1}(-9)\oplus\\
    346 L_{\hat{A_1}}^{w_0}(-9)\oplus 530 L_{\hat{A_1}}^{w_0}(-10)\oplus
    482 L_{\hat{A_1}}^{w_1}(-10)\oplus 714 L_{\hat{A_1}}^{w_1}(-11)\oplus\\
    797 L_{\hat{A_1}}^{w_0}(-11)\oplus 1080 L_{\hat{A_1}}^{w_1}(-12)\oplus
    1180 L_{\hat{A_1}}^{w_0}(-12)
\end{multline}
This result can be expressed using the power series expansion of the branching functions \cite{kac1990idl}.
\begin{eqnarray}
  \label{eq:29}
  \begin{array}{cc}
    b^{(w_1)}_{0}= & 1 + 4\,q^{1}+ 8\,q^{2}+ 15\,q^{3}+ 29\,q^{4}+ 51\,q^{5}+ 85\,q^{6}+ 139\,q^{7}+\\
     &222\,q^{8}+ 346\,q^{9}+ 530\,q^{10}+ 797\,q^{11}+ 1180\,q^{12}+\dots\\
  \end{array}\\
  \begin{array}{cc}
    b^{(w_1)}_{1}= &2+2\,q^{1}+8\,q^{2}+12\,q^{3}+26\,q^{4}+42\,q^{5}+78\,q^{6}+120\,q^{7}+\\
    & 202\,q^{8}+306\,q^{9}+482\,q^{10}+714\,q^{11}+1080\,q^{12}+\dots
  \end{array}
\end{eqnarray}
Here the lower index of the branching function denotes the number of the corresponding $\hat{A_1}$ fundamental weight $w_0=\lambda_0=(0,1,0),\; w_1=\alpha/2=(1,1,0)$.

\section{Physical applications}
\label{sec:phys-appl}
Here we want to discuss possible applications of the described techniques in physical models.

@@

There exist three level one dominant weights $[1,0,0],\;[0,1,0],\;[0,0,1]$ in the weight space of $\hat{A}_2$.
It is easy to see that the set $\Delta_{\bot}$ is empty.

Here we view the representation with the highest weight
$w_0=(1,0,0)$ in detail.

The set of the anomalous points of this representation up to sixth grade is depicted in the Figure \ref{fig:affine_A2_anom_point}. We also show the root subspace of the subalgebra $\hat{A}_1$ under consideration.

\begin{figure}[h!tb]
  \includegraphics[width=160mm]{AffineA2_A1_anom.pdf}
  \caption{The anomalous points of the $(1,0,1,0)$ representation of the algebra $\hat{A_2}$}
  \label{fig:affine_A2_anom_point}
\end{figure}

As the next step of our algorithm \ref{sec:algorithm} we project the anomalous points to the weight space of the subalgebra $\hat{A_1}$ and calculate the dimensions of the corresponding $\mathfrak{a}_{\bot}$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$.
The result of this computation up to the twelfth grade is presented at the Figure \ref{fig:AffineA2_A1_anom_proj}
\begin{figure}[h!tb]
  \centering
  \includegraphics[width=150mm]{AffineA2_A1_proj_anom.pdf}
  \caption{Projected anomalous points and the dimensions of $\mathfrak{a}_{\bot}$-modules.}
  \label{fig:AffineA2_A1_anom_proj}
\end{figure}

Using the definition (\ref{fan-defined}) we get the following set of
the points $\Gamma$ with the corresponding values $s(\gamma+\gamma_0)$, depicted at the Figure \ref{fig:AffineA2A1Fan}.
Here we restricted the computation to the twelfth grade.
\begin{figure}[p]
  \centering
  \includegraphics[width=130mm]{AffineA2_A1_fan.pdf}
  \caption{Fan for $\hat{A_1}\subset \hat{A_2}$}
  \label{fig:AffineA2A1Fan}
\end{figure}

Using the recurrent relation for the anomalous branching coefficients we get the following result
\begin{figure}[p]
  \centering
  \includegraphics[width=130mm]{AffineA2_A1_branching.pdf}
  \caption{Anomalous branching coefficients for $\hat{A_1}\subset \hat{A_2}$}
  \label{fig:AffineA2_A1_branching}
\end{figure}

We see that only two anomalous branching coefficients inside the main
Weyl chamber of $\hat{A}_1$ are non-zero and equal to the branching
coefficients. So the finite reducibility theorem holds and we get
\begin{equation}
  \label{eq:43}
  L^{[1,0,0]}_{\hat{A_2}\downarrow \hat{A_1}}= L_{\hat{A_1}}^{[0,4]}\oplus L_{\hat{A_1}}^{[4,0]}
\end{equation}

For the other level one dominant weights of $\hat{A}_2$ we get trivial
branching rules
\begin{eqnarray}
  \label{eq:44}
   L^{[0,1,0]}_{\hat{A_2}\downarrow \hat{A_1}}= L_{\hat{A_1}}^{[2,2]}\\
   L^{[0,0,1]}_{\hat{A_2}\downarrow \hat{A_1}}= L_{\hat{A_1}}^{[2,2]}
\end{eqnarray}

Using this result we can construct modular-invariant partition function
\begin{equation}
  \label{eq:45}
  Z=\left|\chi_{[4,0]}+\chi_{[0,4]}\right|^2+2\chi_{[2,2]}^2
\end{equation}
